
```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggmosaic)
library(dplyr)
library(epitools)
library(epiR)
```

```{r, echo = FALSE, message = FALSE}
library(LaF)
```

```{r, echo = FALSE}
widths <- read.csv("Extracted pages from 2021_YRBS_Data_Users_Guide_508.xlsx - Sheet2.csv", header=TRUE)
```

```{r, echo = FALSE}
dat0 <- laf_open_fwf("XXH2021_YRBS_Data.dat", column_widths=widths$Length, column_names=widths$Variable, column_types=rep("string", nrow(widths)))
```

```{r, echo = FALSE}
dat <- dat0[,]; str(dat)
```

```{r, includingSubsetOfVariables} 
v2k <- c('Q1',	'Q2',	'Q3',	'Q4',	'RACEETH',	'Q8', 'QN8',	'Q9', 'QN9', 'Q10', 'QN10',	'Q11', 'QN11',	'Q12', 'QN12',	'Q13', 'QN13',	'Q14', 'QN14',	'Q15', 'QN15',	'Q16', 'QN16',	'Q17', 'QN17',	'Q18', 'QN18',	'Q19', 'QN19',	'Q20', 'QN20',	'Q21',	'Q22', 'QN22',	'Q23', 'QN23',	'Q24', 'QN24',	'Q25', 'QN25',	'Q26', 'QN26',	'Q27',	'Q28', 'QN28',	'Q29',	'Q30', 'QN30',	'Q31', 'QN31',	'Q32', 'QN32',	'Q33', 'QN33',	'Q34', 'QN34',	'Q35',	'Q36',	'Q37', 'QN37',	'Q38', 'QN38',	'Q40', 'QN40',	'Q41', 'QN41',	'Q42', 'QN42',	'Q43',	'Q44', 'QN44',	'Q45', 'QN45',	'Q46', 'QN46',	'Q47',	'Q48',	'Q49', 'QN49',	'Q50', 'QN50',	'Q51',	'Q52', 'QN52',	'Q53', 'QN53',	'Q54',	'Q55', 'QN55',	'Q56', 'QN56',	'Q57', 'QN57',	'Q58', 'QN58',	'Q59', 'QN59',	'Q60',	'Q61',	'Q62', 'QN62',	'Q63', 'QN63',	'Q64',	'Q65',	'Q68',	'Q69', 'QN69',	'Q70',	'Q71',	'Q72',	'Q73', 'QN73',	'Q74',	'Q76',	'Q77', 'QN77',	'Q78',	'Q79',	'Q80',	'Q82', 'QN82',	'Q83', 'QN83',	'Q84', 'QN84',	'Q85', 'QN85', 'Q86',	'Q87', 'QN87',	'Q88', 'QN88',	'Q89',	'Q93', 'QN93',	'Q95',	'Q96', 'QN96', 'QN97', 'QN99', 'Q99')
```

```{r}
'%!in%' <- Negate('%in%') #not in function. shows which variables are not in the dataset. should be 0 
```

```{r}
v2k[which(v2k %!in% names(dat))]
```

```{r}
sdat <- dat[,v2k] 
```

```{r}
sdat$Q2 <- factor(sdat$Q2, levels=rev(1:2), labels=rev(c("Female", "Male")))
sdat$QN85 <- factor(sdat$QN85, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$Q4 <- factor(sdat$Q4, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN8 <- factor(sdat$QN8, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN9 <- factor(sdat$QN9, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN10 <- factor(sdat$QN10, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN11 <- factor(sdat$QN11, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN12 <- factor(sdat$QN12, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN13 <- factor(sdat$QN13, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN14 <- factor(sdat$QN14, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN15 <- factor(sdat$QN15, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN16 <- factor(sdat$QN16, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN17 <- factor(sdat$QN17, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN18 <- factor(sdat$QN18, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN19 <- factor(sdat$QN19, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN20 <- factor(sdat$QN20, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN22 <- factor(sdat$QN22, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN23 <- factor(sdat$QN23, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN24 <- factor(sdat$QN24, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN25 <- factor(sdat$QN25, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN26 <- factor(sdat$QN26, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN28 <- factor(sdat$QN28, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN30 <- factor(sdat$QN30, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN31 <- factor(sdat$QN31, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN32 <- factor(sdat$QN32, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN33 <- factor(sdat$QN33, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN34 <- factor(sdat$QN34, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN37 <- factor(sdat$QN37, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN38 <- factor(sdat$QN38, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN40 <- factor(sdat$QN40, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN41 <- factor(sdat$QN41, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN42 <- factor(sdat$QN42, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN44 <- factor(sdat$QN44, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN45 <- factor(sdat$QN45, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN46 <- factor(sdat$QN46, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN49 <- factor(sdat$QN49, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN50 <- factor(sdat$QN50, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN52 <- factor(sdat$QN52, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN53 <- factor(sdat$QN53, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN55 <- factor(sdat$QN55, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN56 <- factor(sdat$QN56, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN57 <- factor(sdat$QN57, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN58 <- factor(sdat$QN58, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN59 <- factor(sdat$QN59, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN62 <- factor(sdat$QN62, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN69 <- factor(sdat$QN69, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN73 <- factor(sdat$QN73, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN77 <- factor(sdat$QN77, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN82 <- factor(sdat$QN82, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN83 <- factor(sdat$QN83, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN84 <- factor(sdat$QN84, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN87 <- factor(sdat$QN87, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN88 <- factor(sdat$QN88, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN93 <- factor(sdat$QN93, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN96 <- factor(sdat$QN96, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN97 <- factor(sdat$QN97, levels=rev(1:2), labels=rev(c("Yes", "No")))
sdat$QN99 <- factor(sdat$QN99, levels=rev(1:2), labels=rev(c("Yes", "No")))


```

```{r}
tab <- table(sdat$Q2); tab
ptab <- prop.table(tab); ptab
barplot(ptab*100)
```

```{r}
table(sdat$Q2, sdat$QN85, useNA="ifany")
tab <- table(sdat$Q2, sdat$QN85) 
ptab <- prop.table(tab); ptab
```

Based on the contigency table, females reported having mental health that wasn't good the majority of the time as opposed to males. 



```{r}
ggplot(data = sdat, mapping = aes(x = Q65)) +
  geom_bar() + 
  labs(title = "Sexual Identity among YRBS Students", #1=Heterosexual(straight), 2=GayOrLesbian, 3=Bisexual
       x = "Sexual Identity",                         #
       y = NULL)
```


```{r, NewQ65Variable}
sdat <- sdat %>%
  mutate(sexual_identity = case_when(Q65 == 1 ~ "Heterosexual",
                                 	Q65 == 2 ~ "Gay or Lesbian",
                                 	Q65 == 3 ~ "Bisexual",
                                 	Q65 == 4 ~ "Other Identity",
                                 	Q65 == 5 ~ "Other Identity",
                                  TRUE ~ NA))
```

```{r, DichotomousSexIdentVariable}
sdat$sexual_identity2 <- factor(ifelse(sdat$sexual_identity == "Heterosexual", "Heterosexual", "Sexual Minority"))
#This creates a dichotomous sexual identity variable 
```

```{r}
table(sdat$sexual_identity)
```

```{r}
table(sdat$sexual_identity2)
```




Although the first code chunk gave each response for Q65 a category/sexual identity, the second dichotomous variable will facilitate the analyses. Since some categories don't have enough data/very few data, there's an issue when running the analyses and the creation of a dichotomous variable allows for fewer and broader categories.


```{r, CategorizingSexualIdentity}
ggplot(data = sdat, mapping = aes(x = sexual_identity)) + #put in descending order?? #add color??
  geom_bar() +        #add quantity to bars??
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, color = "black")
  labs(title = "Sexual Identity among YRBS Students", 
       x = "Sexual Identity",
       y = NULL)
```
```{r}
sdat %>%
  count(sexual_identity)
```



Most students within this dataset identity as Heterosexual (straight). 



```{r, CategorizingGradeLevel}
sdat <- sdat %>%
  mutate(grade_level = case_when(Q3 == 1 ~ "9th",
                                 Q3 == 2 ~ "10th",
                                 Q3 == 3 ~ "12th",
                                 TRUE ~ "Other"))
```

```{r, GradeDistribution}
ggplot(data = sdat, mapping = aes(x = grade_level)) + #put into descending order
  geom_bar() + 
  labs(title = "Grade Level Counts in YRBS",
       x = "Grade Level",
       y = NULL)
```

The grade level with most participants within the YRBS are 9th graders. 

```{r}
sdat <- sdat %>% 
  mutate(raceneth = case_when(
    RACEETH == 1 ~ "Indian/Alaska Native",
    RACEETH == 2 ~ "Asian",
    RACEETH == 3 ~ "Black/African American",
    RACEETH == 4 ~ "Native Hawaiian/Other PI",
    RACEETH == 5 ~ "White",
    RACEETH == 6 ~ "Hispanic/Latino",
    RACEETH == 7 ~ "Hispanic/Latino",
    RACEETH == 8 ~ "Other",
    TRUE ~ "Other"
  ))
```

```{r}
sdat$raceneth2 <- factor(ifelse(sdat$raceneth == "White", "White", ifelse(sdat$raceneth == "Black/African American", "Black/AA", ifelse(sdat$raceneth == "Hispanic/Latino", "Hispanic/Latino", "Other"))))
```

```{r} 
sdat$raceneth2 <- relevel(sdat$raceneth2, ref="White") #Code makes White individuals the reference (base-line)
```

The code above creates a new variable with 4 levels as opposed to 7 levels. Like variable Q65, having more levels results in fewer outcomes for each level at the time of analyses which causes troubles in the analyses. Fewer categories allows for larger outcomes in each category. 

```{r, RaceEthnicity}
ggplot(data = sdat, mapping = aes(y = raceneth)) + 
  geom_bar() + 
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, color = "black")
  labs(x = NULL,
       y = NULL,
       title = "Race/Ethnicity Distribution in YRBS")
```

Students of white race are the most common within the YRBS. 

```{r}
ggplot(data = sdat) +
  geom_mosaic(aes(x = product(QN83),
                  fill = sexual_identity)) +
  labs(title = "Students Tested for STDs Other than HIV in Past 12 Months",
       y = "Sexual Identity")
  
```

```{r}
tab <- table(sdat$QN83); tab
ptab <- prop.table(tab); ptab
barplot(ptab*100)
```

94.34% of students were not tested for STDS other than hIV within the last 12 months. 

```{r}
table(sdat$Q2, sdat$QN83, useNA="ifany")
tab <- table(sdat$Q2, sdat$QN83) 
ptab <- prop.table(tab); ptab
```


```{r}
ggplot(data = sdat, mapping = aes(x = QN82)) + 
  geom_bar()
```

```{r, data frame w/ hetersexual individuals only}
heterosexualdata <- sdat %>%
  filter(sexual_identity == "Heterosexual")
```

```{r, echo = FALSE, results='hide'}
print(heterosexualdata)
```

```{r}
heterosexualdata %>%
  count(raceneth, QN10)
```

```{r, TextingWHileDriving}
heterosexualdata %>%
  count(raceneth, QN11)
```

```{r, CarriedGunToSchool}
heterosexualdata %>%
  count(raceneth, QN13)
```

```{r, AttemptedSuicide}
heterosexualdata %>%
  count(raceneth, QN28)
```

```{r, ScreenedForSTIs}
heterosexualdata %>%
  count(raceneth, QN83)
```

```{r, CigaretteUsage}
heterosexualdata %>%
  count(raceneth, QN30)
```

```{r, CurrentAlcoholUsers}
heterosexualdata %>%
  count(raceneth, QN41)
```

```{r, SexualPartner#>4}
heterosexualdata %>%
  count(raceneth, QN59)
```

```{r, MentalHealthReports}
heterosexualdata %>%
  count(raceneth, QN85)
```

```{r, PhysicalActivity}
heterosexualdata %>%
  count(raceneth, QN77)
```

```{r, FruitIntake}
heterosexualdata %>%
  count(raceneth, QN69)
```

```{r, VegetableIntake}
heterosexualdata %>%
  count(raceneth, QN73)
```

```{r, SleepingAtGuardian/Parent}
heterosexualdata %>%
  count(raceneth, QN87)
```

```{r, WitnessedViolenceNeighborhood}
heterosexualdata %>%
  count(raceneth, QN18)
```

```{r, ParentsAwareOfWhereabouts}
heterosexualdata %>%
  count(raceneth, QN97)
```


```{r, data frame w/ minority individuals only}
sexualminoritydata<- sdat %>%
  filter(sexual_identity != "Heterosexual")
```

```{r, echo = FALSE, results='hide'}
print(sexualminoritydata)
```

```{r}
sexualminoritydata %>%
  count(raceneth)
```

```{r, DrunkDriving}
sexualminoritydata %>%
  count(raceneth, QN10)
```

```{r, TextedWhileDriving}
sexualminoritydata %>%
  count(raceneth, QN11)
```

```{r, CarriedGunSchool}
sexualminoritydata %>%
  count(raceneth, QN13)
```

```{r, AttSuicide}
sexualminoritydata %>%
  count(raceneth, QN28)
```

```{r, ScreenedSTIs}
sexualminoritydata %>%
  count(raceneth, QN83)
```

```{r, CigConsumption}
sexualminoritydata %>%
  count(raceneth, QN30)
```

```{r, CurrentAlcUser}
sexualminoritydata %>%
  count(raceneth, QN41)
```

```{r, #SexualPartners>4}
sexualminoritydata %>%
  count(raceneth, QN59)
```

```{r, mentalHealth}
sexualminoritydata %>%
  count(raceneth, QN85)
```

```{r, PhysHealth}
sexualminoritydata %>%
  count(raceneth, QN77)
```

```{r, FrtIntake}
sexualminoritydata %>%
  count(raceneth, QN69)
```

```{r, VegIntake}
sexualminoritydata %>%
  count(raceneth, QN73)
```

```{r, SleepingAtGuardian }
sexualminoritydata %>%
  count(raceneth, QN87)
```

```{r, WitnessedViolNeighborhood}
sexualminoritydata %>%
  count(raceneth, QN18)
```

```{r, Parent/GuardianAwareOfWhereabout}
sexualminoritydata %>%
  count(raceneth, QN97)
```

##Analyses 

```{r}
ovars <- c("QN10", "QN11", "QN13", "QN28", "QN83", "QN30", "QN41", "QN59", "QN77", "QN69", "QN85", "QN87", "QN18", "QN97")
```

```{r}
results <- rep(list(), length=length(ovars)); names(results) <- ovars
for(i in 1:length(ovars)){
  tab2 <- table(sdat$sexual_identity2, sdat[,ovars[i]], sdat$raceneth2) 
  mhors <- epi.2by2(tab2, method="cross.sectional")
  crude.ors <- mhors$massoc.detail$OR.strata.wald 
  rownames(crude.ors) <- levels(sdat$raceneth2)  
  mh.breslow.day <- mhors$massoc.detail$OR.homog.brday 
  mhadj.or <- mhors$massoc.detail$OR.mh.wald 
  overall.or <- mhors$massoc.detail$OR.crude.wald 
  results[[i]] <- list('Crude Stratified ORs'=crude.ors, 'Breslow Day Test'=mh.breslow.day, 'M-H adjusted OR'=mhadj.or, 'Overall OR'=overall.or)
}
results
```


Through the code above, we're able to obtain the M-H Adjusted ORs, Overall ORs, Breslow Day Tests, and Crude Stratified ORs. I used the results from the Breslow Day test(s) to determine if I should report either the M-H Adjusted OR or the Crude Stratifed OR for each outcome of interest. I recorded the results in a [spreadsheet](https://docs.google.com/spreadsheets/d/1fhsCy5VcsFbPUJEsjyiVMxHLLEcGsuXwiwfcx_ZFglQ/edit?gid=433396617#gid=433396617) where the variables are divided by different outcomes of interest such as risky behaviors, health outcomes, and environmental factors. 

## Creating New Data Frames

In the following code, I'll be extracting the Crude Stratified ORs and the M-H Adjusted ORs from the 'results' list from the previous code chunk. I'll create a data frame for each that'll allow me to use ggplot2 for visualizations. 


```{r, ExtractingCrudeStratifiedORs}
#initialize empty data frame
crude_or_data <- data.frame( #creates data frame w/ columns
  Variable = character(), #stores the name if the variable from 'ovars' list and is initlizaed as a character (avoids R from treating as factor/category)
  Strata = character(), #stores the name of the different strata ("White", "Hispanic", etc)
  OR = numeric(),     #stores ORs as numeric vector
  CI_lower = numeric(), 
  CI_upper = numeric(),
  stringsAsFactors = FALSE #ensures chr vectors aren't automatically turned to factors 
)

# Loop through 'results' to extract crude stratified ORs
for(i in 1:length(results)) { #loop goes through each item/variable in the 'results' list
  crude_ors <- results[[i]]$`Crude Stratified ORs` #pulls out ORs and their conf. intervals
  strata_names <- rownames(crude_ors)   #pulls the names of the different strata that the ORs are calculated for
  
  for(j in 1:nrow(crude_ors)) {   #loop goes through each group within the current variable and pulls out OR for that group
    crude_or_data <- rbind(crude_or_data, data.frame( #adds a new row to table for each group, rbind allows rws to be added together 
      Variable = names(results)[i], #adds names of current variable (like "QN10")
      Strata = strata_names[j],  #adds name of the group (like "Black/AA")
      OR = crude_ors[j, "est"],  #extract the OR
      CI_lower = crude_ors[j, "lower"],  #lower CI
      CI_upper = crude_ors[j, "upper"]   #upper CI
    ))
  }
}

# View the data frame
print(crude_or_data)

```

```{r}
# Initialize empty data frame for M-H adjusted ORs
mhadj_data <- data.frame(
  Variable = character(), 
  OR = numeric(),         
  CI_lower = numeric(),   
  CI_upper = numeric(),   
  stringsAsFactors = FALSE 
)

# Loop through 'results' to extract M-H adjusted ORs and CIs
for(i in 1:length(results)) { #loop goes through each item/variable in the 'results' list
  mhadj_or_data <- results[[i]]$`M-H adjusted OR` #pulls out the M-H adjusted ORs+CIs for the current variable
  
  # Check if mhadj_or_data contains the expected values
  if(nrow(mhadj_or_data) == 1) {  # Ensure it's a single row data frame
    mhadj_data <- rbind(mhadj_data, data.frame(
      Variable = names(results)[i],  #name of the current variable
      OR = mhadj_or_data$est,  #extract OR from "est"
      CI_lower = mhadj_or_data$lower,  #extract lower CI bound
      CI_upper = mhadj_or_data$upper   #extract upper CI bound
    ))
  } else {
    # Warn if data does not match expected format
    warning(paste("Data for", names(results)[i], "is not in the expected format."))
  }
}

# View data frame
print(mhadj_data)

```





## Ploting Crude Stratfied Odd Ratios

Question of interest: Does the effect of sexual identity on health outcomes differ across racial identities?

```{r}
ggplot(crude_or_data, aes(x = Variable, y = OR)) +
  geom_point(size = 3, color = "black") +  #points for OR
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width = 0.2, color = "blue") +  #error bars for CI
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  #line at OR = 1
  coord_flip() +
  labs(x = "Variable",
       y = "Crude Odds Ratio (OR)", 
       title = "Crude Odds Ratios with 95% Confidence Intervals") +
  theme_bw() 

```

```{r}
ggplot(crude_or_data, aes(x = Variable, y = OR, color = Strata, shape = Strata)) +
  geom_point(size = 3) +  #points for OR
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width = 0.7) +  #error bars for CI
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  #line at OR = 1
  facet_wrap(~ Strata) +
  coord_flip() +  #flip coordinates to horizontal plot
  labs(x = "Outcome(s) of Interest", 
       y = "Crude Odds Ratio (OR)", 
       title = "Crude Stratified Odds Ratios by Racial Identity") +
  theme_linedraw() + 
  theme(legend.position = "none")

```

```{r}
crude_selected <- crude_or_data %>% #creates data frame only with those variables I reported Crude ORs for
  filter(Variable %in% c("QN10", "QN41", "QN85", "QN77", "QN69", "QN97")) #selecting desired variables 

crude_selected %>%
ggplot(aes(x = Variable, y = OR, color = Strata, shape = Strata)) +
  geom_point(size = 2.5) +  
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width = 0.4) +  #ymin/ymax specifies upper and lower limits of confidence intervals, width = end line width, errorbar = adds error bars to each point rep 95% CI
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") + #horizontal dashed line @ y=1
  facet_wrap(~ Strata) + #breaks it into diff race/ethnicity/strata
    coord_flip() +  #changes to horizontal (better visual)
  labs(x = "Outcome(s) of Interest",
       y = "Crude Odds Ratio (OR)", 
       title = "Crude Stratified Odds Ratios by Racial Identity") +
  theme_linedraw() +
  theme(legend.position = "none") +
  geom_text(aes(label = round(OR, 2)),  #text labels w/ OR values rounded to 2 decimal places
            vjust = -0.5, #move text slightly above points
            size = 3, #text/label size
            color = "black")
```


For the outcomes of interest above, I reported the Stratified Crude ORs because the Breslow-Day test revealed significant differences (with p-values < 0.05) in how sexual identity affects the outcome across different racial groups. Reflecting on the Stratified Crude ORs will provide a clear view of how these effects vary between groups. The Mantel-Haenszel adjusted ORs, which give an overall average, might not accurately reflect these variations and could lead to ignoring important differences. 

**QN10 Interpretation:**
Black/AA and Hispanic/Latino sexual minority students demonstrate a higher likelihood of drunk driving compared to White heterosexual students. Black/AA sexual minority students show about 2.54 times the odds of drunk driving, and Hispanic/Latino sexual minority students show about 1.64 odds, which is slightly elevated, however not as much as Black/African American students.

**QN41 Interpretation:**
QN41 represents the percentage of students who drank alcohol (at least one drink) on at least one day during the 30 days before the survey. Compared to their White heterosexual counterparts, all three Strata (Black/AA, Hispanic/Latino and Other Racial groups) demonstrate significantly higher odds of drinking.  
- Black/AA sexual minority students show about 2.31 higher odds of drinking alcohol compared to White heterosexual students
- Hispanic/Latino sexual minority students show about 1.50 higher odds of drinking alcohol compared to White heterosexual students
- Other Racial/Ethnic sexual minority groups (including Pacific Islanders and Native Americans) show roughly 1.62 higher odds of drinking compared to White heterosexual students

**QN85 Interpretation**
QN85 regards students who reported that their mental health was not good (including stress, anxiety, and depression) most of the time or always during the 30 days before the survey. For all three strata (Black/AA, Hispanic/Latino, and Other racial/ethnic groups) there is a statistically significant association with reporting poor mental health, as demonstrated by their confidence intervals not crossing 1 (1 represents no difference at all).
- Hispanic/Latino sexual minority students 

**QN97 Interpretation**
QN97 represents the percentage of students who reported that their parents or other adults in their family most of the time or always know where they are going or with whom they will be. The crude stratified ORs for Black/AA, Hispanic/Latino and other racial/ethnic groups show no strong evidence/statistical significance that these strata differ from White students (since the confidence interval for each includes 1). However, when looking at the plots above, it seems that White students are slightly less likely to report that their parent(s)/guardian always know whom they're with. 


## Plotting Mantel-Haenszel Adjusted Odd Ratios

```{r}
mhadj_selected <- mhadj_data %>%
  filter(Variable %in% c("QN11", "QN13", "QN28", "QN83", "QN30", "QN59", "QN87", "QN18"))

mhadj_selected %>%
  ggplot(aes(x = Variable, y = OR)) +
  geom_point(size = 4, color = "blue") +  
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width = 0.5, color = "blue") +  #error bars for CI
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  #line at OR = 1
  coord_flip() +  #flip coordinates to make horizontal
  labs(x = "Outcome(s) of Interest",
       y = "Mantel-Haenszel Adjusted Odds Ratio (OR)", 
       title = "Mantel-Haenszel Adjusted Odds Ratios with 95% Confidence Intervals") +
  theme_bw() + 
   geom_text(aes(label = round(OR, 2)),  #text labels w/ OR values rounded to 2 decimal places
            vjust = -1, #move text slightly above points
            size = 3, #text/label size
            color = "black")
  
```

For the outcomes above, I reported and plotted the Mantel-Haenszel (M-H) Adjusted ORs because the Breslow Day test yielded p-values greater than 0.05 for each, indicating no statistical significance between sexual identity and race/ethnicity on these outcomes. 

Reporting the M-H Adjusted ORs provided measures of the overall effect of sexual identity identity on these outcomes while accounting for race/ethnicity. This approach combines data across all ethnic/racial groups and yields a single measure (OR) (with lower and upper bound confidence intervals), giving a clearer picture when the effect doesn't change significantly between groups. 















